---
title: "Session 5"
author: "Tolga Akturk"
date: '2022-03-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Continuing on from session 4 

### Data Wrangling

```{r}

survey<-read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-18/survey.csv")

country<-survey%>%
  distinct(country)

survey%>%
  count(country, sort=TRUE)

survey_short<-survey%>%
  filter(
    country=="United States" |
      country=="USA" |
      country=="US" |
      country=="U.S." |
      country=="Usa" |
      country=="United States of America"
    )

survey_short%>%
  filter(
    annual_salary>50000 &
      gender=="Woman"
  )

survey_short%>%
  select(how_old_are_you, industry, annual_salary, gender)%>%
  filter(gender=="Woman")%>%
  count(industry, sort=TRUE)

knitr::kable(head(mtcars[, 1:4]), "pipe")

``` 

## Session 5 starts here

Arranging observations

```{r}

survey_short%>%arrange(desc(annual_salary))%>%
  select(industry, annual_salary)

survey_short%>%count(industry, sort=TRUE)

```


Groupings and New Variables 

```{r}

survey_short%>%
  mutate(total_salary=annual_salary+other_monetary_comp)%>%
  select(industry, total_salary, annual_salary, other_monetary_comp)

survey_short%>%
  mutate(average_salary=mean(annual_salary, na.rm=TRUE))%>%
  select(industry, average_salary)

survey_short%>%
  group_by(industry)%>%
  mutate(average_salary=mean(annual_salary, na.rm=TRUE))%>%
  select(industry, average_salary)

survey_short%>%
  group_by(industry, gender)%>%
  mutate(mean_salary=mean(annual_salary, na.rm=TRUE))%>%
  select(industry, gender, mean_salary)%>%
  arrange(desc(mean_salary))%>%
  ungroup()



#Add a new variable max_salary which is maximum salary per industry
  
```


Summarising Data

```{r}

survey_short%>%
  group_by(industry)%>%
  summarise(average_salary=mean(annual_salary, na.rm=TRUE))

#Group by industry and gender. Use summarise for average_salary
#Group by gender and industry. Use summarise for average_salary
#Group by industry and gender. Use summarise for average_salary and max_salary

survey_short%>%
  group_by(industry, gender)%>%
  summarise(average_salary=mean(annual_salary, na.rm=TRUE))

survey_short%>%
  group_by(gender, industry)%>%
  summarise(average_salary=mean(annual_salary, na.rm=TRUE))

survey_short%>%
  group_by(industry, gender)%>%
  summarise(
    average_salary=mean(annual_salary, na.rm=TRUE),
    max_salary=max(annual_salary, na.rm=TRUE))


```

Case_when

```{r}

survery_short_veryshort<-survey_short%>%
  select(industry, gender, annual_salary)

#if annual_salary is more than 700000 then your annual_salary is high otherwise it is low

survery_short_veryshort%>%
  mutate(
   annual_salary_level=case_when(
     annual_salary>700000 ~"high",
     T~"low"
   )
  )%>%
  arrange(desc(annual_salary))

#if annual_salary is more than 700000 then your annual_salary is high and if your annual_salary is more  than 70000 then your annual_salary is middle otherwise it is low

survery_short_veryshort%>%
  mutate(
   annual_salary_level=case_when(
     annual_salary>700000 ~ "high",
     annual_salary>70000 & annual_salary<=700000 ~ "middle",
     T~"low"
   )
  )

#create a variable highly_paid_women
#if gender=="woman" and annual_salary>50000 then (~) it is "high", otherwise it is low

survery_short_veryshort%>%
  mutate(
   highly_paid_women=case_when(
     gender=="Woman" & annual_salary>50000 ~ "high",
     T~"low"
   )
  )

```


## Session 5 cont. plus Session 6

Notes: 

Must have! 
1. Data is data
2. Aesthethics is variables e.g. x and y axis
3. Geometries are charts e.g. bar, histogram

Optional
4. Facets are seperate graphs and how you position them together
5. Statistics e.g. mean, etc
6. Coordinates e.g. legend, label on coordinates
7. Theme e.g. colour, 

x axis first in code then y


```{r}

#Experimenting with Aesthetics
data(mpg)
mpg%>%ggplot( 
  aes(displ, hwy, colour = class, size = cyl, shape = trans)) +
geom_point() 

#one continutous, one discrete
data(mpg)
mpg%>%ggplot( 
  aes(manufacturer, cty, colour = class)) +
geom_point() 

#two continuous

data(mpg)
mpg%>%ggplot( 
  aes(hwy, cty)) +
geom_smooth() 

``` 

```{r}

data(mpg)
mpg%>%ggplot( 
  aes(displ, hwy, colour = class)) +
geom_point() 

```


```{r}
#Step 11 assignment part 2
#install.packages("skimr")
library(skimr)

skim(survey)

#generate as many as possible beautiful and meaningful data viz using survey datasheet

```

```{r}

survey%>%ggplot( 
  aes(industry, annual_salary )) +
geom_boxplot() 

```
